---
title: "Vitamine C increases the tooth growth in ginea pigs"
author: "Juan Sebastián Beleño Díaz"
date: "31 de enero de 2016"
output: 
    pdf_document: default
    html_document:
        keep_md: yes
---

## Overview

This paper shows important discoveries about the implication of vitamine C in
tooth growth in the population of ginea pigs, to do this job was necessary a 
use of basic concepts in exploratory data analysis and statistic inference.

## Basic exploratory analysis

In this section it'll be shown some metadata related to the TowthGrowth dataset 
that will be used to perform some analysis in this project, before starting the 
project is very useful to know about the size of the dataset, its rows and columns, 
its format, etc.

```{r}
# Load the libraries needed and ToothGrowth data
library(datasets)
library(ggplot2)
data(ToothGrowth)

# Quick look at the dataset variables, the size of the dataset, and the type of 
# each variable, and some of the values that each variable contains
str(ToothGrowth)
```

As we can see in this part the ToothGrowth dataset is a data.frame that has 3 
variables(`len`, `supp` and `dose`), `dose` and `len` are numeric variables and
`supp` is a factor variable, the complete dataset has 60 observations, so it 
means we won't have problems with RAM while doing the processing part. There is
a common patern in the `dose` variable that suggest that maybe there is a few
bunch of values that could take, for that reason it'll be converted this
variable to factor.

```{r}
# dose variable conversion to factor
ToothGrowth$dose <- as.factor(ToothGrowth$dose)

# Quick look of the dataset to analyze if exist a reduced amount of values for
# dose, which is quite probable
str(ToothGrowth)
```

Here we found the first discovery, just exist 3 values for `dose` (0.5, 1 and 2) 

## Summary of the data

According to the official documentation available in
<https://stat.ethz.ch/R-manual/R-devel/library/datasets/html/ToothGrowth.html> 
. *The response is the length of odontoblasts (cells responsible for tooth growth)
in 60 guinea pigs. Each animal received one of three dose levels of vitamin C 
(0.5, 1, and 2 mg/day) by one of two delivery methods, (orange juice or ascorbic 
acid (a form of vitamin C and coded as VC).*

Now that we have some information about the dataset we'll obtain a summary of the
dataset to analyze the results and see what we can do with that data.

```{r}
summary(ToothGrowth)
```

Here we can see that the `len` variable could have values from 4.20 to 33.90 in
this sample, the mean is 18.81 and the median 19.25, this data is given for the
complete dataset but these `len` values could change taking just one value of
`supp` or `dose`.

```{r}
# To see the number of values per combination between dose and supp
table(ToothGrowth$supp, ToothGrowth$dose)
```

Indeed we found 10 `len` values for each combination between `supp` and `dose`. 

```{r, fig.cap="len values when dose and supp change"}
g <- ggplot(aes(x = dose, y = len), data = ToothGrowth) 
g <- g + geom_boxplot(aes(fill = dose)) + facet_wrap(~ supp)
g
```

In the graphic we can see (without a real proof) that more Vitamine C highly 
contributes to the tooth growth, and Orange Juice (OJ) seems to be a better 
presentation in terms of tooth growth than the Ascorbic Acid (VC).

## Hypothesis about tooth growth by supp and dose

The last paragraph in the last section are the hypothesis that will be
tested in this part, so first starting by setting $H_0$ and $H_A$.

$H_0$: Vitamine C doesn't contribute to tooth growth

$H_A$: Vitamine C does contribute to tooth growth

To try to prove our hypothesis is necessary to set comparison in Vitamine
C levels between 0.5 and 1.0, 1.0 and 2.0 and finally 0.5 and 2.0.

```{r}
# Comparison between 0.5 and 1.0
t.test(len ~ dose, data = subset (ToothGrowth, dose %in% c(0.5, 1.0)))
```

```{r}
# Comparison between 1.0 and 2.0
t.test(len ~ dose, data = subset (ToothGrowth, dose %in% c(1.0, 2.0)))
```

```{r}
# Comparison between 0.5 and 2.0
t.test(len ~ dose, data = subset (ToothGrowth, dose %in% c(0.5, 2.0)))
```

These results are in favor of $H_A$ with 95% of confidence, the p-values are 
always under 0.05 and the confidence interval doesn't contain 0.

Having this result let's attack the next hypothesis.

$H_0$: Orange Juice and Ascorbic Acid contribute equal to tooth growth

$H_A$: Orange Juice and Ascorbic Acid doesn't contribute equal to tooth growth

```{r}
# Comparison between Orange Juice and Ascorbic Acid
t.test(len ~ supp, data = ToothGrowth)
```

This result indicates that we can't reject null hypotesis due to p-values is
over 0.05 and the confidence interval contains 0.

## Conclusions

* Vitamine C highly contributes to increase the tooth length in guinea pigs
* There is no difference between give Orange Juice and Ascorbic Acid in the 
tooth growth in guinea pigs

# Assumptions
* All the ginea pigs were the same age and eat a similar food
* The sample of ginea pigs was random in the general population of ginea pigs
* The sample is representative of genral population if ginea pigs